\documentclass[a4paper,oneside]{memoir}
\input{preamble}

\title{Scholar: A Tool for Structured Literature Searches}
\author{Daniel Bosk \and Dan-Claude van Thropic}
\date{\today}

\begin{document}
\frontmatter
\maketitle

\begin{abstract}
Systematic literature reviews are fundamental to rigorous scholarship, yet
conducting them remains tedious and error-prone.
Researchers must search multiple databases with different interfaces, merge
and deduplicate results, and meticulously document their screening decisions.
Scholar addresses these challenges through a command-line tool that unifies
access to five major academic databases---Semantic Scholar, OpenAlex, DBLP,
Web of Science, and IEEE Xplore---while automating the bookkeeping that
systematic reviews demand.

The tool centers on a terminal-based review interface where researchers
navigate papers, read abstracts, record keep/discard decisions with
justifications, and organize findings with themes and notes.
For large result sets, an LLM-assisted classification system learns from
user decisions to suggest classifications for pending papers, dramatically
reducing manual screening effort while maintaining human oversight.
Intelligent caching, automatic deduplication, and PDF management round out
the workflow support.

This document is the complete literate source of Scholar, weaving together
the design rationale, implementation, and tests into a unified narrative.
The code is extracted (``tangled'') from this document using noweb.
\end{abstract}
\vfill
\inputminted[linenos=false]{text}{../LICENSE}

\clearpage
\tableofcontents*

\mainmatter

\input{introduction}

\part{Core Functionality}

\input{../src/scholar/scholar.tex}
\input{../src/scholar/cache.tex}

\input{../src/scholar/providers.tex}
\input{../src/scholar/enrich.tex}
\input{../src/scholar/pdf.tex}

\part{Review System}

\input{../src/scholar/notes.tex}
\input{../src/scholar/review.tex}
\input{../src/scholar/llm_review.tex}

\part{User Interface}

\input{../src/scholar/cli.tex}
\input{../src/scholar/tui.tex}

\backmatter

\end{document}
